\label[overflow]
\app Additional graphs

\sec Additional graphs

\label[ap_dataset_longedgesplit]
\secc Splitting long edges
%------------------

Figures \ref[36_splitedge_valerr_img] and \ref[36_splitedge_boxplot_img] show additional graphs to the experiment in \ref[ch6_dataset_longedgesplit].

\midinsert
		\clabel[36_splitedge_valerr_img]{Split long edges, Image model, validation error over iterations}
		\picw=9cm \cinspic graphs/6.1.2._splitting_long_edges-minlen/fig2_average_valerr_img.pdf
		\caption/f Model: \textit{Image}, Datasets: \textit{original}, \textit{minlen10}, \textit{minlen20}, \textit{minlen30}. Averages of validation errors with edge splitting experiment.
\endinsert

\midinsert
		\clabel[36_splitedge_boxplot_img]{Split long edges, Image model, comparison of last epoch between datasets}
		\picw=9cm \cinspic graphs/6.1.2._splitting_long_edges-minlen/fig1_boxplot_last_img.pdf
		\caption/f Model: \textit{Image}, Datasets: \textit{original}, \textit{minlen10}, \textit{minlen20}, \textit{minlen30}. Last epoch of the edge splitting experiment.
\endinsert


%\midinsert
%\line{\hsize=.5\hsize \vtop{%
%		\clabel[36_splitedge_valerr_img]{Split long edges, Image model, validation error over iterations}
%		\picw=7.5cm \cinspic graphs/6.1.2._splitting_long_edges-minlen/fig2_average_valerr_img.pdf
%		\caption/f Averages of validation errors with Image model testing edge splitting.
%   \par}\vtop{%
%		\clabel[36_splitedge_boxplot_img]{Split long edges, Image model, comparison of last epoch between datasets}
%		\picw=7.5cm \cinspic graphs/6.1.2._splitting_long_edges-minlen/fig1_boxplot_last_img.pdf
%		\caption/f Last epoch of edge splitting with Image model.
%   \par}}
%%\centerline {aaa\hfil\hfil bbb}\nobreak\medskip
%\endinsert


\supereject

\label[ap_model_base]
\secc Different CNN base model for feature transfer
%------------------

\midinsert
\clabel[ap_base_models]{Different base CNN on 299x299 dataset}
\picw=17cm \cinspic graphs/6.2.1._different_CNN/___manual_out_299.pdf
\caption/f Model: \textit{Mixed with variable base}, Dataset: \textit{5556x\_mark\_res\_299x299}. Illustration of five models and their averages of validation and training errors. Note that this dataset uses images of 299x299 pixel size.
\endinsert

\label[ap_model_versus] 
\secc Model competition - Image vs. OSM vs. Mixed
%------------------

Figures \ref[ap1_modelversus_evol], \ref[ap1_modelversus_last] and \ref[ap1_modelversus_best] provide additional graphs for the section \ref[ch6_model_versus]. Here we can see the "5556x_minlen30_640px" dataset.

\midinsert
\clabel[ap1_modelversus_evol]{Model competition, evolution over epochs}
\picw=12cm \cinspic graphs/6.2.2._model_competition-img-vs-osm-vs-mix/fig1_evolution_minlen30.pdf
\caption/f Models: \textit{OSM}, \textit{Image}, \textit{Mixed}, Dataset: \textit{5556x\_minlen30\_640px}. Model competition, evolution over epochs.
\endinsert

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[ap1_modelversus_last]{Model competition, last epoch}
      \picw=5.5cm \cinspic graphs/6.2.2._model_competition-img-vs-osm-vs-mix/fig2_last_epoch_minlen30.pdf
      \caption/f Model competition, last epoch. Dataset: \textit{5556x\_minlen30\_640px}.
   \par}\vtop{%
      \clabel[ap1_modelversus_best]{Model competition, best epoch}
      \picw=5.5cm \cinspic graphs/6.2.2._model_competition-img-vs-osm-vs-mix/fig3_best_epoch_minlen30.pdf
      \caption/f Model competition, best epoch. Dataset: \textit{5556x\_minlen30\_640px}.
   \par}}
%\centerline {aaa\hfil\hfil bbb}\nobreak\medskip
\endinsert

%\supereject

\label[ap_model_osm_wd]
\secc OSM specific - width and depth
%------------------

Figure \ref[ap_osm_wd_d2cut] adds to the experiment of \ref[ch6_model_osm_wd]

\midinsert
\clabel[ap_osm_wd_d2cut]{OSM model with fixed depth and variable width}
\picw=12cm \cinspic graphs/6.2.3._OSM_model_GRID_4x4/fig2_cutOfD3fixed.pdf
\caption/f Model: \textit{OSM}, Dataset: \textit{5556x\_markable\_640x640}. OSM model with depth fixed at value 3 and variable width.
\endinsert

\endinput