\label[overflow]
\app Dataset overview

\label[ap_osm_vec_attrval]
\sec OSM vector details
%------------------

Table \ref[ap_attrval_pairs] shows us the list of attribute and value pairs, which build up the OSM vectors. We are interested only in the occurances of the pairs in this list. Note that we show only a start of the vector, which is in total made of 594 attribute-value pairs. Refer to section \ref[ch3_osm_vector] for details.

\midinsert
\clabel[ap_attrval_pairs]{OSM attribute - value pairs}
\picw=16cm \cinspic tables/ap._sql_attr_val_combos.pdf
\captiontablehack/f Examples of attribute=value pairs used to filter out which values are of interest to us while building the OSM vector.
\endinsert


\label[ap_dataset_list]
\sec Used datasets overview
%------------------

Table \ref[ap_used_datasets] shows the list of the used datasets. These can be roughly divided by their purpose and method of generation. We can see those created by the edge splitting technique from \ref[ch6_dataset_longedgesplit] and those generated with data augmentation schemes from \ref[ch6_dataset_augmentation]. We also have several datasets used purely for fast prototyping and experimentation, which can serve as simple examples.

\midinsert
\clabel[ap_used_datasets]{List of used datasets}
\picw=15cm \cinspic tables/ap_datasets_list.pdf
\captiontablehack/f List of used datasets divided by their use.
\endinsert

\label[ap_dataset_examples]
\sec Dataset examples

\secc Original dataset
%------------------

Original dataset "5556x_markable_640x640" containing images of pixel size 640x640 in Figure \ref[ap_original].

\midinsert
\clabel[ap_original]{Original dataset examples}
\picw=12cm \cinspic tables/tab_original_dataset.pdf
\caption/f Scored examples of the dataset "5556x_markable_640x640". Note that the score labels come from the initial dataset.
\endinsert

\secc Expanded dataset
%------------------

Expanded dataset "5556x_markable_640x640_2x_expanded" with images generated expanding the original dataset using the first augmentation scheme of \ref[code7b_data_aug] in Figure \ref[ap_extended].

\midinsert
\clabel[ap_extended]{Extended dataset examples}
\picw=12cm \cinspic tables/tab_expanded_dataset.pdf
\caption/f Examples of the dataset "5556x_markable_640x640_2x_expanded".
\endinsert

\secc Aggressively extended dataset
%------------------

Expanded dataset "5556x_markable_640x640_2x_agressive_expanded" with images generated expanding the original dataset using the second, more aggressive augmentation scheme of \ref[code7b_data_aug] in Figure \ref[ap_ag_extended].

\midinsert
\clabel[ap_ag_extended]{Aggressively extended dataset examples}
\picw=12cm \cinspic tables/tab_ag_expanded_dataset.pdf
\caption/f Examples of the dataset "5556x_markable_640x640_2x_agressive" "_expanded".
\endinsert

\label[ap_dataset_histo]
\sec Dataset analysis
%------------------

We look at the composition of dataset entries in terms of edge length before their splitting in Figures \ref[ap_edgelen_histo], \ref[ap_edgelen_boxplot] and \ref[ap_edgelen_sorted]. We look at the score distribution in Figures \ref[ap_scores_histo], \ref[ap_scores_boxplot] and \ref[ap_scores_sorted].

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[ap_edgelen_histo]{Edge length analysis, histogram}
      \picw=8cm \cinspic figures/dataset_stats/edge_lengths-histo.pdf
      \caption/f Histogram of edge lengths from the initial dataset \textit{5556x\_markable} \textit{\_640x640}.
   \par}\vtop{%
      \clabel[ap_scores_histo]{Score distribution analysis, histogram}
      \picw=8cm \cinspic figures/dataset_stats/score-hist.pdf
      \caption/f Histogram of image scores from the initial dataset \textit{5556x\_markable} \textit{\_640x640}.
   \par}}
\endinsert


\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[ap_edgelen_boxplot]{Edge length analysis, box plot}
      \picw=8cm \cinspic figures/dataset_stats/edge_lengths-boxplot.pdf
      \caption/f Edge lengths in the initial dataset \textit{5556x\_markable\_640x640}. We split this dataset in experiment \ref[ch6_dataset_longedgesplit].
   \par}\vtop{%
      \clabel[ap_scores_boxplot]{Score distribution analysis, box plot}
      \picw=8cm \cinspic figures/dataset_stats/score-boxplot.pdf
      \caption/f Images score distribution in the initial dataset \textit{5556x\_markable} \textit{\_640x640}.
   \par}}
\endinsert

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[ap_edgelen_sorted]{Edge length analysis, sorted values}
      \picw=8cm \cinspic figures/dataset_stats/edge_lengths-plot.pdf
      \caption/f Edge lengths from the initial dataset \textit{5556x\_markable\_640x640} sorted by value.
   \par}\vtop{%
      \clabel[ap_scores_sorted]{Score distribution analysis, sorted values}
      \picw=8cm \cinspic figures/dataset_stats/score-plot.pdf
      \caption/f Image scores from the initial dataset \textit{5556x\_markable\_640x640} sorted by value. Note that we are using images generated from valid edges, which had score labels in the initial dataset.
   \par}}
\endinsert


\endinput